---
title: "HW4"
author: "Akhil Repaka | ar78978"
date: "`r Sys.Date()` | https://github.com/akhilrepaka123/HW4---SDS-315/blob/main/HW4.Rmd"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)
library(mosaic)
library(ggplot2)
library(knitr)
text <- readLines("brown_sentences.txt", warn = FALSE)
letter_freq <- read.csv("letter_frequencies.csv", stringsAsFactors = FALSE)
```


# Problem 1

```{r}
# problem 1
set.seed(123)

sim_fraud = do(100000) * nflip(n=2021, prob=0.024) # monte-carlo simulations of the null hypothesis 

ggplot(sim_fraud) + 
  geom_histogram(aes(x=nflip), binwidth=1) + geom_vline(xintercept = 70, color = 'red', linetype = 'dashed') #distribution for the monte -  carlo simulations and plotting where the test statistic is

total_fraud = sum(sim_fraud >= 70) # how many test cases are equal to or above the test statistic

p_value_fraud = total_fraud/100000 
```


**Null Hypothesis:** Iron Bank's 70 flagged securities trades are consistent with the SEC’s 2.4% baseline rate and therefore other traders in the long run.

**Test Statistic:** This is the number of observed flagged cases, which is 70. The P-Value is determined using the simulations that result in a number higher than the test statistic. 

**P-Value:** The P_value is **`r p_value_fraud`**

**Conclusion:** Since the p-value is lower than 0.05, this allows us to reject the null hypothesis and assume there was suspicious trading activity going on. 

\newpage

# Problem 2
```{r}
# problem 2
set.seed(123)

sim_restaurant = do(100000) * nflip(n=50, prob=0.03)
ggplot(sim_restaurant) + 
  geom_histogram(aes(x=nflip), binwidth=1) + geom_vline(xintercept = 8, color = 'red', linetype = 'dashed')

total_restaurant = sum(sim_restaurant >= 8)

p_value_restaurant = total_restaurant/100000
```

**Null Hypothesis:** Gourmet Bites is consistent with the Health Department’s 3% baseline rate for average restaurants cited health code violations. 

**Test Statistic:** This is the number of observed flagged cases, which is 8. The P-Value is determined using the simulations that result in a number higher than the test statistic. 

**P-Value:** The P_value is **`r p_value_restaurant`**

**Conclusion:** Since the p-value is lower than 0.05, this allows us to reject the null hypothesis and assume there was a significant reason behind the 8 helath code violations and is not arbritary. 

\newpage
# Problem 3
```{r}
# problem 3
set.seed(123)

expected_props <- c(0.30, 0.25, 0.20, 0.15, 0.10)  # Population distribution
observed_count <- c(85, 56, 59, 27, 13)  # Empaneled jurors
total_jurors <- sum(observed_count)

expected_count <- total_jurors * expected_props

chisq <- sum((observed_count - expected_count)^2 / expected_count) # manually calculating for chi square value

chi_sq_test <- chisq.test(observed_count, p = expected_props) # chi squared function and shows p - value

chi_sq_test
```

Manually calculated observed chi squared value is **`r chisq`**

I used the **chi squared test** to determine whether the distribution of jurors empaneled by this judge is
significantly different from the county’s population proportions. 

**Null Hypothesis:** The judge’s empaneled jurors are selected randomly according to the county's demographic proportions.

**Test Statistic:** The test statistic is calculated using the chi-square function and is **12.426**

**P-Value:** The p-value is 0.01445 or in other words, there is only a 1.44% chance that this observed distribution happens.

**Conclusion:** Since the p-value is lower than 0.05 and the observed chi-square value is higher than the critical chi-square value (~9.49) the distribution is significant and we can reject the null hypothesis.

\newpage
# Problem 4a
```{r}
# problem 4a

clean_sentences = gsub("[^A-Za-z]", "", text)  # Remove non-letters
clean_sentences = toupper(clean_sentences)

calculate_chi_squared = function(text, freq_table) 
  {
  
  # Ensure letter frequencies are normalized and sum to 1
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  # Remove non-letters and convert to uppercase
  clean_sentence = gsub("[^A-Za-z]", "", text)
  clean_sentence = toupper(clean_sentence)
  
  # Count the occurrences of each letter in the sentence
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
  # Calculate expected counts
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  # Chi-squared statistic
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}

chi_squared_values = sapply(clean_sentences, calculate_chi_squared, freq_table = letter_freq)

hist(chi_squared_values, main = "Reference Distribution of Chi-Squared Statistics",
     xlab = "Chi-Squared Value")

```



# Problem 4b
```{r}
#problem 4b

newsentences <- c(
 "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
 "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
 "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
 "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
 "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
"The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
 "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
 "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product #that exceeded everyone’s expectations."
)
newsentences <- data.frame(text = newsentences)

for(i in 1:length(newsentences$text)) {
  newsentences$chi_squared_val[i] = calculate_chi_squared(newsentences$text[i], letter_freq)
  newsentences$p_val[i] <- 1 - round(pchisq(newsentences$chi_squared_val[i], df = 25), 3) # 25 degrees of freedom
}
kable(newsentences, booktabs = TRUE, longtable = TRUE,
      col.names = c("Sentence", "Chi-Squared Value", "P-Value"),
      caption = "Chi-Squared Values and P-Values for Sentences")
```

